Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-09 ms/data_readme.Rmd
Update data readme

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_dryad_file",
    "comm_ferns_dryad_file",
    "shape_ferns_dryad_file",
    "fern_traits_dryad_file",
    "japan_fern_phylogram_dryad_file",
    "japan_fern_tree_dryad_file",
    "japan_fern_tree_uncollapsed_dryad_file"
    )
  )
```

This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')`

@@

- _targets.tar.gz: Tarball (compressed folder) including all workflow results produced by R targets package
- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_deer_range.gpkg: Distribution maps of Japanese deer (Cervus nippon) in Japan
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of native, non-hybrid ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for native, non-hybrid ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_traits_lucid.csv: Morphological trait data of native, non-hybrid ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a labeled map of Japan
- japan_map.gpkg: Map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Selected analysis results. Files include:
  - japan_ferns_biodiv.csv: Biodiversity statistics of native, non-hybrid ferns in Japan
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_shape.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs collapsed to polytomies
  - japan_ferns_tree_uncollapsed.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs not collapsed
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan

Additional data used in analysis not included in current data package: Data 
including DNA sequences, taxonomy, and breeding system data for native Japanese 
ferns included in the Dryad data package of @Ebihara2019b (https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32)
were used in the analysis.

@@

ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and
phylogenetic tree (Nitta et al., in prep.). For more information, see README
file contained in zipped archive.

MD5 checksum: `r tools::md5sum(here::here("data/ftol_data_release_v0.0.1.zip"))`

@@

```{r occ-point-data-summary-cols, results = "asis"}
  occ_point_data_summary %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens; 'occ_point_data_ferns' indicates data only includes specimens after filtering",
      "variable", "Variable. 'n_taxa' indicates number of taxa, 'n_specimens' indicates number of specimens",
      "value", "Value corresponding to each variable")
  )
```

@@

Categorical trait codes as follows:
- 0 = absent
- 1 = common
- 2 = rare
- 3 = uncertain
- 4 = common and misinterpreted
- 5 = rare and misinterpreted
- 6 = not scoped

@@

`r fs::path_file(japan_map_file)`: Map of land area of Japan. Raw data downloaded from
Geospatial Information Authority of Japan under the Creative Commons Attribution
License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html). All polygons
were combined into a single polygon and CRS set to JGD2000.

@@

2021-09-07

>>>>

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ebihara_2019_zip_file",
    "ftol_zip_file",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_dryad_file",
    "comm_ferns_dryad_file",
    "shape_ferns_dryad_file",
    "fern_traits_dryad_file",
    "japan_fern_phylogram_dryad_file",
    "japan_fern_tree_dryad_file",
    "japan_fern_tree_uncollapsed_dryad_file",
    "refs_yaml"
    )
  )
```

This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')` by Joel Nitta

@@

- _targets.tar.gz: Tarball (compressed folder) including all workflow results produced by R targets package
- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_deer_range.gpkg: Distribution maps of Japanese deer (Cervus nippon) in Japan
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of native, non-hybrid ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for native, non-hybrid ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_traits_lucid.csv: Morphological trait data of native, non-hybrid ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a labeled map of Japan
- japan_map.gpkg: Map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Selected analysis results. Files include:
  - japan_ferns_biodiv.csv: Biodiversity statistics of native, non-hybrid ferns in Japan
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_shape.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs collapsed to polytomies
  - japan_ferns_tree_uncollapsed.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs not collapsed
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan

Additional data used in analysis not included in current data package:

- doi_10.5061_dryad.4362p32__v4.zip: Data including DNA sequences, taxonomy, and 
breeding system data for native Japanese ferns [@Ebihara2019b] downloaded from 
https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32. 
MD5 checksum `r tools::md5sum(here::here(ebihara_2019_zip_file))`
- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment 
and phylogenetic tree (Nitta et al., in prep.) downloaded from 
https://figshare.com/s/87b8ec102303bf9f3212.
MD5 checksum `r tools::md5sum(here::here(ftol_zip_file))`
FIXME: replace with DOI when available

@@

@@

```{r occ-point-data-summary-cols, results = "asis"}
  occ_point_data_summary %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens, 'occ_point_data_ferns' indicates data only includes specimens after filtering",
      "variable", "Variable. 'n_taxa' indicates number of taxa, 'n_specimens' indicates number of specimens",
      "value", "Value corresponding to each variable")
  )
```

@@

Categorical trait codes as follows:

- 0 = absent
- 1 = common
- 2 = rare
- 3 = uncertain
- 4 = common and misinterpreted
- 5 = rare and misinterpreted
- 6 = not scoped

@@

`r fs::path_file(japan_map_file)`: Map of land area of Japan. Raw data downloaded from
Geospatial Information Authority of Japan under the Creative Commons Attribution
License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html). All polygons
were combined into a single polygon and CRS set to JGD2000 using code available at
https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/R/process_raw_data.R.

@@

2021-09-09
